---
title: "Intriguing Visualizations of Data from World Happiness Reports"
description: |
  
  *A data science project, created by data scientist Dhruv Syngol, to visualize happiness data from a wide spectrum of countries*
  
  Below are some exciting visualizations that represent data collected from World Happiness Reports.
 
  Link to Source of Data: https://www.kaggle.com/ajaypalsinghlo/world-happiness-report-2021?select=world-happiness-report-2021.csv
  
  Check out the *About* page to learn more about the data and the project. 
  
  *NOTE: The graphs are best viewed in landscape mode!*

site: distill::distill_website
---

```{r setup, echo = FALSE}
library(tidyverse)
library(dplyr)
library(plotly)
library(gridExtra)
library(ggpubr)

```

```{r fig.height = 14, fig.width = 14, echo = FALSE}
world_happiness <- read_csv("~/Downloads/world-happiness-report.csv")
world_2021 <- read_csv("~/Downloads/world-happiness-report-2021.csv")

```

```{r, echo = FALSE}
x1 <- world_2021%>%
  rename(country = 'Country name', region = 'Regional indicator', ladder = 'Ladder score', gdp = 'Logged GDP per capita', life_expec = 'Healthy life expectancy')%>%
  select(country, region, ladder, gdp, life_expec)%>%
  group_by(region)%>%
  mutate(`Life Ladder` = mean(ladder), `GDP Per Capita` = mean(gdp), `Life Expectancy` = mean(life_expec))%>%
  ggplot(aes(x=`GDP Per Capita`, y =fct_reorder(region, `GDP Per Capita`))) + 
  geom_col(position = "dodge") + 
    theme(text = element_text(size = 15)) +

  labs(title = "Mean GDP Per Capita by Region", subtitle = "Western Europe, North America and ANZ are both high in Mean GDP Per Capita", x = "Logged GDP Per Capita", y = "Mean Values", caption = "Data from World Happiness Report 2021") 
```

```{r, echo = FALSE}
y1 <- world_2021%>%
  rename(country = 'Country name', region = 'Regional indicator', ladder = 'Ladder score', gdp = 'Logged GDP per capita', life_expec = 'Healthy life expectancy')%>%
  select(country, region, ladder, gdp, life_expec)%>%
  group_by(region)%>%
  mutate(`Life Ladder` = mean(ladder), `GDP Per Capita` = mean(gdp), `Life Expectancy` = mean(life_expec))%>%
  ggplot(aes(x=`Life Expectancy`, y =fct_reorder(region, `Life Expectancy`))) + 
  geom_col(position = "dodge") + 
    theme(text = element_text(size = 15)) +
  labs(title = "Mean Healthy Life Expectancy by Region", subtitle = "Western Europe, North America, and ANZ lead World Regions", x = "Healthy Life Expectancy", y = "Region", caption = "Data from World Happiness Report 2021") 

```

```{r, echo = FALSE}
z1 <- world_2021%>%
  rename(country = 'Country name', region = 'Regional indicator', ladder = 'Ladder score', gdp = 'Logged GDP per capita', life_expec = 'Healthy life expectancy')%>%
  select(country, region, ladder, gdp, life_expec)%>%
  group_by(region)%>%
  mutate(`Life Ladder` = mean(ladder), `GDP Per Capita` = mean(gdp), `Life Expectancy` = mean(life_expec))%>%
  ggplot(aes(x=`Life Ladder`, y =fct_reorder(region, `Life Ladder`))) + 
  geom_col(position = "dodge") + 
  theme(text = element_text(size = 15)) +
  labs(title = "Mean Life Ladder Scores by Region", subtitle = "North America and ANZ Leads All Regions", x = "Life Ladder Scores", y = "Region", caption = "Data from World Happiness Report 2021") 

```
```{r fig.height = 15, fig.width = 12, echo = FALSE}
ggarrange(x1, y1, z1, nrow = 3)
```


<aside>
We can see that the regions of "North America and ANZ" and "Western Europe" dominate in mean life ladder scores, healthy life expectancy, and GDP per capita. 

The regions of "Sub-Saharan Africa" and "South Asia" score low in the following statistics. 

</aside>


For the following graphs, move your cursor on a line in the interactive visualizations to see the year, life ladder score, and country name.

```{r fig.height = 11, fig.width = 12, echo = FALSE}
data_first <- read.csv("~/Downloads/archive/world-happiness-report-2021.csv")
data_second <- read.csv("~/Downloads/archive/world-happiness-report.csv")
data_first <- data_first%>%select(Country.name, Regional.indicator)
merged <- merge(data_first, data_second, by = "Country.name", all = TRUE)
first_merge <- merged%>%
  group_by(Country.name)%>%
  arrange(Country.name)%>%
  na.omit(merged)%>%
  filter(year >= 2010 & year <= 2019)%>%
  filter(Regional.indicator %in% c("Southeast Asia", "Western Europe", "South Asia", "Commonwealth of Independent States"))
y <- ggplot(data = first_merge, mapping = aes(x = year, y = Life.Ladder, color = Country.name)) + 
  geom_line() + 
  facet_wrap(~Regional.indicator) + 
  scale_x_continuous(breaks = 2010:2019) +
  theme(legend.position = 'none', strip.background = element_rect(colour = "black", fill = "white"), text = element_text(size=10)) +
  labs(title = "Trends in Life Ladder Scores in 2010-2019", subtitle = "Western European Countries Have Higher Life Ladder Scores", x = "Years (2010-2019)", y = "Life Ladder Score", caption = "Data from the World Happiness Report")
ggplotly(y)
```


<aside>



In this graph, we can see that Western European countries have higher life ladder scores than countries that comprise the Commonwealth of Independent States, South Asia, and Southeast Asia. 

We see an upward trend in Life Ladder scores in the countries of the Commonwealth of Independent States; however, the trend is inverse for South Asia. 

Southeast Asia and Western Europe seem to be stable in their Life Ladder scores over the 10 years. 

</aside>



```{r fig.height = 11, fig.width = 12, echo = FALSE}
second_merge <- merged%>%
  group_by(Country.name)%>%
  arrange(Country.name)%>%
  na.omit(merged)%>%
  filter(year >= 2010 & year <= 2019)%>%
  filter(Regional.indicator %in% c("Southeast Asia", "Western Europe", "South Asia", "Commonwealth of Independent States"))

y <- ggplot(data = second_merge, mapping = aes(x = year, y = Log.GDP.per.capita, color = Country.name)) + 
  geom_line() + 
  facet_wrap(~Regional.indicator) + 
  scale_x_continuous(breaks = 2010:2019) +
  theme(legend.position = 'none', strip.background = element_rect(colour = "black", fill = "white"), text = element_text(size=10)) +
  labs(title = "Trends in GDP Per Capita (Natural Log) in 2010-2019", subtitle = "Western European Countries have a Higher GDP Per Capita", x = "Years (2010-2019)", y = "GDP Per Capita (Nautral Log)", caption = "Data from the World Happiness Report")
ggplotly(y)
```

<aside>


In this graph, we can see an upward trend in GDP per capita in all regions highlighted in the graph. 

Western European countries tend to have a higher GDP per capita.  
</aside>


```{r fig.height = 11, fig.width = 12, echo = FALSE}
third_merge <- merged%>%
  group_by(Country.name)%>%
  arrange(Country.name)%>%
  na.omit(merged)%>%
  filter(year >= 2010 & year <= 2019)%>%
  filter(Regional.indicator %in% c("Southeast Asia", "Western Europe", "South Asia", "Commonwealth of Independent States"))
y <- ggplot(data = third_merge, mapping = aes(x = year, y = Healthy.life.expectancy.at.birth, color = Country.name)) + 
  geom_line() + 
  facet_wrap(~Regional.indicator) + 
  scale_x_continuous(breaks = 2010:2019) +
  theme(legend.position = 'none', strip.background = element_rect(colour = "black", fill = "white"), text = element_text(size=10)) +
  labs(title = "Trends in Healthy Life Expectancy at Birth in 2010-2019", subtitle = "Western European Countries have Higher Healthy Life Expectancy at Birth", x = "Years (2010-2019)", y = "Healthy Life Expectancy at Birth", caption = "Data from the World Happiness Report")
ggplotly(y)

```

<aside>


Similar to the GDP per capita graph, there are upward trends in the regions highlighted for the healthy life expectancy at birth statistic. 

Western European countries tend to have high, yet similar healthy life expectancy at birth statistics. 

We can also see that countries like Nepal and Kazakhstan have made significant jumps in healthy life expectancy. 

</aside>

For the following graphs, move your cursor on a point in the interactive visualizations to GDP per capita, life ladder score year, region, and country name.

Double-click on a region in the legend to isolate specific points of that region on the graph. You can click on other regions to compare countries in different regions and their respective scores. 

Double-click on the graph if you want to zoom out.

In the first graph, I wanted to create a scatter plot with the life ladder scores and GDP per capita on one graph. I plotted each country and filled each point by region.  

```{r fig.height = 11, fig.width = 15, echo = FALSE}

fourth_graph <- world_2021%>%
  ggplot(aes(x = `Logged GDP per capita`, y = `Ladder score`, color = `Regional indicator`, text=paste("Country name:",`Country name`))) + 
  geom_jitter() + 
  theme(text = element_text(size=12), legend.text = element_text(size = 10)) + 
  labs(title= "A Comparison of Life Ladder Score and GDP Per Capita (Logged)", x = "GDP Per Capita (Natural Log)", y = "Life Ladder Score", caption = "Data from World Happiness Report 2021") 
ggplotly(fourth_graph)

```

<aside>


In this graph, we can see that countries in Western Europe have high GDP per capitas and high life ladder scores (as seen in the top-right corner of the graph). 

Countries in the bottom-left corner, including those from Sub-Saharan Africa and South Asia, have lower GDP per capitas and lower life ladder scores. 

This graph can help contribute to a conclusion that countries in Western European, North America, as well as Australia and New Zealand, tend to have a better quality of life compared to other countries of the world. 

</aside>

The freedom and social support scores signify the extent to which each statistic impacted beneficial changes in life quality. I wanted to plot these scores for each country and fill each point by region.

```{r fig.height = 11, fig.width = 15, echo = FALSE}

fifth_graph <- world_2021%>%
  ggplot(aes(x = `Freedom to make life choices`, y = `Social support`, color = `Regional indicator`, text=paste(`Country name`))) + 
  geom_jitter() + 
  theme(text = element_text(size=12), legend.text = element_text(size = 10)) + 
  labs(title= "A Comparison of Freedom of Life Choices and Social Support", x = "Freedom to Make Life Choices", y = "Social Support", caption = "Data from World Happiness Report 2021") 
ggplotly(fifth_graph)

```

<aside>


Similar to the graph above, we can see that countries in Western Europe and North America and ANZ have more freedom to make life choices and have access to more social support. 

Countries in the bottom-left corner have less freedom and social support. 

Countries in regions like South Asia are plotted all around the graph and are not strictly clustered like countries in Southeast Asia and North America and ANZ.

</aside>